---
title: "Problem Set 2 - Statistics"
output: html_notebook
---

The purpose of this problem set is to explore some simple statistics concepts with R.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

### Generate a Normal Distribution

R has a number of built-in functions to help you create datasets for testing. For example, you could generate a set of data that represents a Gaussian distribution, sometimes called a "Normal" curve. We generate a set of 100 with a mean of 0 and a standard deviation of 1. Then we plot it as a histogram.

```{r}
set.seed(123)
x = rnorm(n=100, mean = 0, sd = 1)
hist(x,breaks=10)
```

The theoretical probability of points lying within 1 standard deviation of the mean is 0.67. 

### How many points are there within 1 standard deviation of the mean 0?

```{r}
count1sigma = length(which((x < 1) & (x > -1)))
count1sigma                     
```

### Add a Gaussian curve to the histogram
Use the plot, curve, and abline functions to add the normal curve, mean, median, and standard deviations to the plot. Include a legend in the plot.

```{r}
mu = mean(x)
sigma = sd(x)
med = median(x)
hist(x,breaks=15, main="Gaussian Curve over Histogram",prob=T)
curve(dnorm(x, mean=mu, sd=sigma), 
          col="darkblue", lwd=2, add=TRUE, yaxt="n")
abline(v = mu,
 col = "royalblue",
 lwd = 2)

abline(v = med,
 col = "red",
 lwd = 2)
sdupper = mu + sigma
sdlower = mu - sigma
abline(v = sdupper,
 col = "green",
 lwd = 1)

abline(v = sdlower,
 col = "green",
 lwd = 1)
legend(x = "topright", # location of legend within plot area
 c("Mean", "Median","Std Deviation"),
 col = c("royalblue", "red", "green"),
 lwd = c(2, 2, 2))

```
### Explore the Old Faithful dataset
There is a built-in dataset in R called "faithful". Explore some of its properties.

#### What is the dimension of the dataset?

#### What are the two columns?

#### Plot the waiting time. Does the histogram look Gaussian?

#### Plot waiting on the x axis and eruptions (in minutes) on the y axis. What do you observe?

```{r}
plot (faithful$waiting, faithful$eruptions)

```


#### Do you think there is a relationship between erruptions and waiting time? Why or why not? What do you think the relationship might be? Could the relationship be explored with a linear model?

#### Could your observations help build a better model for predicting eruption times when you have already waited for 68 minutes?


### Explore the sample dataset.

To do this, 


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

y <- rt(200, df = 5)
qqnorm(y); qqline(y, col = 2)
qqplot(y, rt(300, df = 5))

dir()

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
