---
title: "Problem Set 2 - Statistics"
output: html_notebook
---

The purpose of this problem set is to explore some simple statistics concepts with R.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

### Generate a Normal Distribution

R has a number of built-in functions to help you create datasets for testing. For example, you could generate a set of data that represents a Gaussian distribution, sometimes called a "Normal" curve. We generate a set of 100 with a mean of 0 and a standard deviation of 1. Then we plot it as a histogram.

```{r}
set.seed(123)
x = rnorm(n=100, mean = 0, sd = 1)
hist(x,breaks=10)
```

The theoretical probability of points lying within 1 standard deviation of the mean is 0.67. 

### How many points are there within 1 standard deviation of the mean 0?

```{r}
count1sigma = length(which((x < 1) & (x > -1)))
count1sigma                     
```

### Add a Gaussian curve to the histogram


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
